### 2022-09-13 src/presentation/presentation.qmd
Add first set of videos to the presentation

---
title: "The value of intermediation"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        transition: fade
bibliography: references.bib
---

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>

# Intorduction

## Motivation

TODO

## Main ideas

 - Use the Shapley-value instead of an extensive-form bargaining model
    - Analytically more tractable
    - Has bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
    - Has precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - Focus on the "few large players -- continuum of small players" case (oceanic games)
    - Good approximation of the finite player case even for not too many players
    - Results have a nice geometric interpretation
 - Balance generality and having useful results

## Preview of results

 - The Shapley-value seems to be a very tractable way of modeling bargaining in certain important contexts
    - One/two-sided platforms
    - Upstream-downstream setups
 - Comparative statistics give support to the bargaining interpretation
 - Extensions to include some important cases are relatively straightforward
 - Even a simple application yields interesting results about hybrid platforms

#

## The Shapley-value

An example game in coalitional form

:::: {.columns}

::: {.column width="50%"}
Imagine a game with three players:

 - A wholesale producer: $P$
 - Two downstream resellers: $A_1, A_2$

:::

::: {.column width="10%"}
:::

::: {.column width="40%"}
And the following achievable profits:

Coalition | Profits
----------|---------
$\{P\},  \{A_i\}$     |  0
$\{A_1, A_2\}$  | 0  
$\{P, A_i\}$  | 2 
$\{P, A_1, A_2\}$  | 3
:::

::::

## The Shapley-value

Each player gets their expected marginal contribution

:::: {.columns}

::: {.column width="50%"}
::: {.r-stack}

::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/draw_objects.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P12.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P21.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_2P1.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_1P2.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_12P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_21P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/final.mp4" />
:::

:::
:::

::: {.column width="10%"}
:::

::: {.column width="40%"}
And the following achievable profits:

Coalition | Profits
----------|---------
$\{P\},  \{A_i\}$     |  0
$\{A_1, A_2\}$  | 0
$\{P, A_i\}$  | 2 
$\{P, A_1, A_2\}$  | 3
:::

::::

## Why do we care about it?

Same idea as with the Nash bargaining game/solution

 - Axiomatic reasoning
    - The Shapley-value is the only sharing rule that is efficient, symmetric, linear, and satisfies the null player axiom [@shapley1953additive]
 - Bargaining foundations
    - The value arises of the subgame-perfect equilibrium of various alternating-offer bargaining games [e.g. @gul1989bargaining; @winter1994demand; @hart1996bargaining; @stole1996intra; @inderst2003bargaining]
 - A tractable and sensible-looking gain-sharing rule
    - Comparative statics show that it behaves as one would expect

# The relevant setting

## Baseline model

# Thank you

## References

::: {#refs}
:::