### 2022-09-15 src/presentation/presentation.qmd
Switch to mathjax 3 for better output

---
title: "The value of intermediation"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        html-math-method: mathjax
        transition: fade
        toc: true
        toc-depth: 1
        toc-title: Agenda
bibliography: references.bib
---

# Intorduction

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>

## Motivation

 - Bargaining between participants is rather overlooked in platform settings
   - Most papers implicitly assume take-it-or-leave-it offers
 - The use of cooperative bargaining concepts is not widespread in IO
   - Although they are well established in certain subfields, such as upstream-downstream markets
 - Most papers only consider finite-player games
   - Modeling the small players as a continuum simplifies analysis

::: {.notes}
 - Relevance to competition policy
 - However: today's presentation is mostly theory but if I have time I'll say a few words about applications
:::

## Main ideas

 - Use the Shapley-value instead of an extensive-form bargaining model
    - Analytically more tractable
    - Has bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
    - Has precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - Focus on the "few large players -- continuum of small players" case (oceanic games)
    - Good approximation of the finite player case even for not too many players
    - Results have a nice geometric interpretation
 - Balance generality and having useful results

## Related literature

 - Shapley-value as a bargaining outcome
   - [@gul1989bargaining; @winter1994demand; @hart1996bargaining; @stole1996intra]
 - Cooperative bargaining in IO models
   - [@billera1978internal; @hart1990property; @levy1997individual; @segal2003collusion; @inderst2003bargaining; @montez2007downstream]
 - Values of oceanic games
   - [@hart1973values; @fogelman1980asymptotic]
 - Hybrid platforms
   - [@anderson2021hybrid; @hagiu2022should]

::: {.notes}
 - Further evidence that the Shapley-value is a good bargaining outcome
 - Cooperative bargaining solutions in IO models are finite-player
 - The asymptotics of the weighted value are unknown
 - The literature on hybrid platforms is very recent
 - Shapley-value in a platform setting is new
:::

## Preview of results

 - The Shapley-value seems to be a very tractable way of modeling bargaining in certain important contexts
    - One/two-sided platforms
    - Upstream-downstream setups
 - Comparative statistics give support to the bargaining interpretation
 - Extensions to include some important cases are relatively straightforward
 - Even a simple application yields interesting results about hybrid platforms

# The Shapley-value

::: {.notes}
 - Just the intuition in a few words
:::

## The Shapley-value

An example game in coalitional form

:::: {.columns}

::: {.column .fragment width="50%"}
Imagine a game with three players:

 - Two producers: $A_1, A_2$
   - can each make \$1 on their own
 - A platform: $P$
   - Triples the firms' profits

:::

::: {.column width="10%"}
:::

::: {.column .fragment width="40%"}
Coalition | Profits
----------|---------
$\{P\}$     |  0
$\{A_i\}$   | 1
$\{A_1, A_2\}$  | 2
$\{P, A_i\}$  | 3
$\{P, A_1, A_2\}$  | 6
:::

::::

::: {.notes}
 - Two ingredients: players (left) and characteristic function (right)
 - We don't care about strategies, only about value achieved by coalitions
:::

## The Shapley-value

Each player gets their expected marginal contribution

:::: {.columns}

::: {.column width="50%"}
::: {.r-stack}

::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/draw_objects.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P12.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P21.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_2P1.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_1P2.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_12P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_21P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/final.mp4" />
:::

:::
:::

::: {.column width="10%"}
:::

::: {.column width="40%"}
Coalition | Profits
----------|---------
$\{P\}$     |  0
$\{A_i\}$   | 1
$\{A_1, A_2\}$  | 2
$\{P, A_i\}$  | 3
$\{P, A_1, A_2\}$  | 6
:::

::::

::: {.notes}
 - Most direct definition of Shapley value
 - Closed-form expression
 - Generally computationally expensive, but not in this case
:::

## Why do we care about it?

Same idea as with the Nash bargaining game/solution

 - Axiomatic reasoning
    - The Shapley-value is the only sharing rule that is efficient, symmetric, linear, and satisfies the null player axiom [@shapley1953additive]
 - Bargaining foundations
    - The value arises of the subgame-perfect equilibrium of various alternating-offer bargaining games [e.g. @gul1989bargaining; @winter1994demand; @hart1996bargaining; @stole1996intra; @inderst2003bargaining]
 - A tractable and sensible-looking gain-sharing rule
    - Comparative statics show that it behaves as one would expect

::: {.notes}
 - You don't need to think about cooperative game theory in the following part
 - Think of it as a convenient gain sharing rule with some (cooperative and non-cooperative) micro-foundations
:::

# Main results

::: {.notes}
 - Very simple setting
 - Conveys the intuition behind the results
:::

## Baseline model

:::: {.columns}

::: {.column width="100%"}

Set of all players: $N = \{P, A_1, \dots, A_n\}$

**Assumption**: the platform is the only way for firms to
reach users and that all $n$ firms are identical
$$
v(C) = \begin{cases}
   0                                           & \text{if } P \notin C \\
   f\left(\frac{|C \setminus \{P\}|}{n}\right) & \text{otherwise}.
\end{cases}
$$
$\forall\, C \subset N$.
:::

::::

::: {.notes}
 - Not only platforms. Also describes
   - upstream-downstream setups
   - generally any setting with an important large players and many identical small ones
:::

## Value of the platform

:::: {.columns}

::: {.column width="40%"}

Characteristic function:
$$
\scriptsize
v(C) = \begin{cases}
   0 \;\text{ if } P \notin C \\
   f\left(\frac{|C \setminus \{P\}|}{n}\right)
\end{cases}
$$

::: {.fragment .fade-in fragment-index=2}
Shapley value:
$$
\scriptsize
\varphi_P^n(f) = \frac{1}{n+1} \sum_{k=0}^{n} f(k)
$$
:::
::: {.fragment .fade-in fragment-index=5}
$$
\scriptsize
\to \int_0^1 f(t) \, \mathrm{d}t
$$
:::

:::

::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/draw_bars.mp4" />
:::
::: {.fragment fragment-index=2}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/add_brace.mp4" />
:::
::: {.fragment fragment-index=3}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/more_firms.mp4" />
:::
::: {.fragment fragment-index=4}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/continuous_approximation.mp4" />
:::
::: {.fragment fragment-index=6}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/surplus_division.mp4" />
:::

:::

:::

::::

::: {.notes}
 - Think of the previous figure with the orderings
 - Main insight: each ordering with the platform at position $k$ has the same probability
:::

## Main proposition

::: {#prp-baseline}
Let $f$ be continuous on $[0, 1]$. Then

$$
\lim_{n \to \infty} \varphi_P^n(f) = \int_0^1 f(t) \, \mathrm{d}t
$$
:::

Even as firms become infinitesimally small, the platform cannot take all the created value

::: {.notes}
 - The bargaining problem is reduced to a simple integral
:::

## Simple comparative statics

:::: {.columns}

::: {.column width="40%"}

::: {.fragment fragment-index=1}
The platform gets a larger slice when
:::
::: {.fragment fragment-index=2}
  1. the firms are substitutes to each other
:::
::: {.fragment fragment-index=3}
  2. the platform has intrinsic value
:::
::: {.fragment fragment-index=4}
  3. the fringe firms have no outside option
:::

:::

::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}
   <video data-autoplay src="../../out/manim_figures/videos/corollary_comparative/800p60/sections/draw_graph.mp4" />
:::
::: {.fragment fragment-index=2}
   <video data-autoplay src="../../out/manim_figures/videos/corollary_comparative/800p60/sections/complements.mp4" />
:::
::: {.fragment fragment-index=3}
   <video data-autoplay src="../../out/manim_figures/videos/corollary_comparative/800p60/sections/intrinsic_value.mp4" />
:::
::: {.fragment fragment-index=4}
   <video data-autoplay src="../../out/manim_figures/videos/corollary_comparative/800p60/sections/outside_option.mp4" />
:::

:::

:::

::::

::: {.notes}
 - The comparative statics seem to coincide with one's intuition about bargaining
:::

# Extensions

::: {.notes}
 - Can be skipped if don't have time
:::

## Multiple platforms

:::: {.columns}

::: {.column width="45%"}
Imagine that

  - There are $m$ platforms
  - They are perfect substitutes
:::

::: {.column width="5%"}
:::

::: {.column .fragment width="45%"}
That is, 
$$
   v(S) = \begin{cases}
      0  \; \text{ if } P_i \notin S \; \forall i \\
      f\left(\frac{\#_A(S)}{n}\right)
   \end{cases}
$$
:::

::::

::: {#prp-multiple-platforms .fragment}
## Multiple platforms
The limit of the Shapley-value (as $n \to \infty$) of each player of type $P$ is
$$
\varphi_{P_i}^{\infty, m} = \int_0^1 (1-t) ^ {m-1} f(t) \mathrm{d}t .
$$
:::

::: {.notes}
 - Probabilistic interpretation: the platform has to be the first one in order to have any additional value
 - As the number of platforms increases, the *total value* they obtain decreases, and the total value firms obtain increases
:::

## Weighted value

::: {.fragment}
 - Endow the game with a weight system
   - The platform has weight $\lambda$
   - The firms have weight $1$
 - Weights can thoughts of as innate bargaining power [@shapley1953additive; @hart1973values]
:::

::: {#prp-weighted .fragment}
## Weighted values
Let $f(t)$ be continuous on $[0, 1]$. Then
$$
\varphi_P^\infty(f, \lambda) = \int_0^1 \underbrace{\lambda t^{\lambda - 1}}_{g(t)} f(t) \mathrm{d}t
$$
:::

::: {.notes}
 - The limit of weighted values is *new*
 - Corollaries give credence to the bargaining power interpretation
 - Probabilistic interpretation (weighting permutations)
:::

## Two-sided bargaining

:::: {.columns}

::: {.column width="45%"}
Imagine that there are two fringe types ($A_i$ and $B_i$)
:::

::: {.column width="5%"}
:::

::: {.column .fragment width="45%"}
$$
    v(S) = \begin{cases}
        0 \; \text{ if } P \notin S \\
        f\left(\frac{\#_A(S)}{n}, \frac{\#_B(S)}{n}\right).
    \end{cases}
$$
:::

::: {#prp-multiple-platforms .fragment}
## Two-sided bargaining
Let $f(a, b)$ continuous on $[0, 1] \times [0, 1]$. Then
$$
f(1, 1) = \underbrace{\int_0^1 f(t, t) \mathrm{d}t}_{\varphi_P^\infty(f)} + \underbrace{\int_0^1 t \frac{\partial f(t, t)}{\partial a} \mathrm{d}t}_{\varphi_A^\infty(f)} + \underbrace{\int_0^1 t \frac{\partial f(t, t)}{\partial b} \mathrm{d}t}_{\varphi_B^\infty(f)}
$$
:::

::::

::: {.notes}
 - The other side (e.g. consumers) also participates in the bargaining process
 - Values are weighted marginal contributions
 - $P$ does not have to be in the "middle" of the transactions
:::

# Example application

::: {.notes}
 - Intentionally simple model
 - Most elements borrowed from [@anderson2021hybrid]
:::

## Market setup

 - Measure 1 of consumers [à la @anderson2021hybrid]
   - Each of them buy a unit of products
   - Logit-type utility function ($u_i = v_i - p_i + \mu \epsilon_i$)
 - Continuum of firms ($a(t), t \in \mathbb{R}$) with one differentiated product each
 - Measure 1 of outside options
 - A platform
   - necessary for the firms to sell their products
   - May or may not have its own products

::: {.notes}
 - Each consumer chooses the products giving them the highest net utility
 - I don't want to go into the details of the model
:::

## Timing of the game

:::: {.columns}

::: {.column width="45%"}
::: {.fragment .fade-in fragment-index=2}
 1. Firms decide whether to invest at cost $c$
:::
::: {.fragment .fade-in fragment-index=3}
 2. Platform entry fees and decisions are negotiated
:::
::: {.fragment .fade-in fragment-index=4}
 3. Firms and platform set product prices, consumers buy products
:::

::: {.fragment .fade-in fragment-index=6}
**Idea:** 2. stage bargaining is over *final allocations*
:::
:::

::: {.column width="55%"}
::: {.r-stack}

::: {.fragment fragment-index=1}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/draw_objects.mp4" />
:::
::: {.fragment fragment-index=2}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/investment.mp4" />
:::
::: {.fragment fragment-index=3}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/platform_entry.mp4" />
:::
::: {.fragment fragment-index=4}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/pricing_game.mp4" />
:::
::: {.fragment fragment-index=5}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/everyone_enters.mp4" />
:::
::: {.fragment fragment-index=6}
<video data-autoplay src="../../out/manim_figures/videos/market_structure/800p60/sections/shapley_values.mp4" />
:::

:::
:::

::::

::: {.notes}
 - Investment decisions are sunk costs
 - Entry fees are lump sum, do not influence pricing
 - All players foresee the outcome of the pricing subgame
:::

## Results

 - Assuming no collusion in the pricing game, total profits as a function of entering firms are
$$
v(N_f) = \frac{N_P B + N_F B}{N_P B + N_F B + 1} \quad \text{ where } B = e^\frac{v - c - \mu}{\mu}
$$
 - If entry fees were fixed, $N_P$ would not influence product variety / consumer welfare
 - However, a higher $N_P$ leads to a better bargaining position for the platform and lower number of entrants

::: {.notes}
 - Constant, additive markup over marginal cost
 - Consumer welfare is a function of product variety
:::

## Results graphically

::: {.r-stack}
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/draw_axes.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/char_func.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/fix_players.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/P_profit_share.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/profit_share_point.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/hybrid_case.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/switch_to_phi_F.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/phi_F_functions.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/cost_function.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/equilibrium.mp4" />
:::
::: {.fragment}
<video data-autoplay src="../../out/manim_figures/videos/equilibrium_entry/800p60/sections/N_F_loss.mp4" />
:::

:::

# Conclusion

## Conclusions

 - Proposed a very tractable way of modeling bargaining applicable to certain important contexts
 - The results are for upstream-downstream situations, platform settings, and two-sided markets
 - The outcomes generally concur with intuitive expectations
 - Even the simple example application provides some interesting insights into hybrid platforms

::: {.notes}
My contributions
 - Shapley-value is a good fit for platform settings
 - The continuous case is vary tractable in this specific setting
 - Results about the weighted value are mine
 - Insight that hybrid platforms have higher bargaining power, leading to more distortion
:::

## Future work

- There is a variety of ways in which the model could be extended (e.g. additional platforms provide some value)
- The example application could be turned into a proper, full-fledged model
- A lab experiment about the outcomes of group bargaining could be conducted
- The model could be applied to other contexts

::: {.notes}
 - Some extensions are straightforward
 - However, some situations just don't fit the model
 - Essentially a plug-and-play approach when one some middle ground between one side taking all the gains
:::

# Thank you {.unnumbered .unlisted}

## References

::: {#refs}
:::