### 2023-04-16 src/presentation/topics_talk.qmd
Start working on topics presentation

---
title: "Hybrid platforms and bargaining"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        html-math-method: mathjax
        transition: fade
        touch: true
        toc: true
        toc-depth: 1
        toc-title: Agenda
        slide-number: c
        progress: true
        controls: false
        theme:
         - default
         - css/custom.scss
        include-in-header:
         - js/mathjax-settings.html
bibliography: references.bib
---

# Introduction

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>