### 2023-04-16 src/presentation/topics_talk.qmd
Add motivation

---
title: "Hybrid platforms and bargaining"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        html-math-method: mathjax
        transition: fade
        touch: true
        toc: true
        toc-depth: 1
        toc-title: Agenda
        slide-number: c
        progress: true
        controls: false
        theme:
         - default
         - css/custom.scss
        include-in-header:
         - js/mathjax-settings.html
bibliography: references.bib
---

# Introduction

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>

## Motivation

::: {.highlight-block}
Examine the bargaining power implications of hybrid platforms and its welfare consequences
:::

::: {.fragment}
 - Hybrid platform
    - Getting more and more common
    - Seemingly obvious concerns and high-profire competition policy cases
    - Comparatively little reserach
:::
::: {.fragment}
 - Bargaining between participants is rather overlooked in platform settings
   - Most papers assume take-it-or-leave-it offers
:::

::: {.notes}

:::

## Main ideas

 - Use the Shapley-value instead of an extensive-form bargaining model
    - Analytically more tractable
    - Has bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
    - Has precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - Abstract away from all frictions
    - Allows me to focus on the bargaining channel
    - Offers a good benchmark model
 - Focus on continuum of small players case for tractability
    - Good approximation of the finite player case even for not too many players
    - Results have a nice geometric interpretation

## References

::: {#refs}
:::