### 2023-04-19 src/presentation/topics_talk.qmd
Add model section to the topics talk

---
title: "Hybrid platforms and bargaining"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        html-math-method: mathjax
        transition: fade
        touch: true
        toc: true
        toc-depth: 1
        toc-title: Agenda
        slide-number: c
        progress: true
        controls: false
        theme:
         - default
         - css/custom.scss
        include-in-header:
         - js/mathjax-settings.html
bibliography: references.bib
---

# Introduction

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>

## Motivation

::: {.highlight-block}
Examine the bargaining power implications of hybrid platforms and its welfare consequences
:::

::: {.fragment}
 - Hybrid platform
    - Getting more and more common
    - Seemingly obvious concerns and high-profire competition policy cases
    - Comparatively little reserach
:::
::: {.fragment}
 - Bargaining between participants is rather overlooked in platform settings
   - Most papers assume take-it-or-leave-it offers
:::

::: {.notes}

:::

## Main ideas

 - Use the Shapley-value instead of an extensive-form bargaining model
    - Analytically more tractable
    - Has bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
    - Has precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - Abstract away from all frictions
    - Allows me to focus on the bargaining channel
    - Offers a good benchmark model
 - Focus on continuum of small players case for tractability
    - Good approximation of the finite player case even for not too many players
    - Results have a nice geometric interpretation

## Preview of results

::: {.fragment}
 - The effects of hybrid platforms depend on the substitutability of the products
   - When they are substitutes, it increases the platform's profit share and decreases the fringe's profits
   - In the presence of an investment cost, it leads to lower fringe entry
   - With Logit-demand, hybrid operation leads to welfare losses
:::
::: {.fragment}
 - This modeling strategy seems promising
   - The model is surprisingly tractable
   - Quite general results can be obtained
:::

# Model

## Players

::: {.fragment}
 - One platform: $P$
   - Without the platform, no value can be created
   - Might have its own products (hybrid mdoe)
:::
::: {.fragment}
 - A continuum of potential fringe entrants: $F_i, i \in \mathbb{R}_0^+$
   - Infinitesimal size
   - Have one product each
   - Can only sell through the platform
:::

## Production

Reduced-form approach: assume some function for total profit:

::: {.fragment}
$$
f(N_P, N_F)
$$
:::

::: {.fragment}
 - $N_P$ is the number of platform products
 - $N_F$ is the number of fringe products/firms
 - Assume that $f$ is increasing in both arguments
:::

## Profit sharing

 - Platform gets $$ \pi_P(N_P, N_F) = \int_0^1 w(s) f(N_P, sN_F) \mathrm{d}s $$
 - Fringe gets the rest

::: {.fragment}
 - Based on the Shapley-value 
   - Bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
   - Precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - $w(s)$ is a weighting function with $\int_0^1 w(s) \mathrm{d}s = 1$
   - Related to the weighted value for a certain class of functions
:::

## Some simple comparative statics

::::: {.columns}

:::: {.column width="40%"}

::: {.fragment fragment-index=1}
 - The platform likes being hybrid
$$
\small
\frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0
$$
:::
::: {.fragment fragment-index=2}
 - It depends™ for the fringe
$$
\begin{align*}
\small\mathrm{sign} \left( \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} \right) = \\
\small\mathrm{sign} \left( \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} \right)
\end{align*}
$$
:::

::::

:::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}

:::
::: {.fragment fragment-index=2}

:::

:::

::::

:::::

## Entry and equilibrium

Now let us endogeneize the number of fringe entrants!

::: {.fragment}
 - Unlimited number of potential entrants
:::
::: {.incremental}
 - Timing
    - Hybrid mode is exogeneously determined $\to N_P$
    - Each fringe firm decides whether to invest at cost $I_F$ $\to N_F$
    - Platform and firms produce and sell products $\to f(N_P, N_F)$
    - Profits are shared according to previous rule $\to \pi_P(N_P, N_F), \pi_F(N_P, N_F)$
:::

## Equilibrium

::: {.incremental}
 - Free entry equilibrium: fringe firms make zero net profits
   $$ \pi_F(N_P, N_F^*) = I_F N_F^* $$
 - In order to guarantee uniqueness, let us assume
   $$
   \scriptsize
   \begin{gather*}
       f'_{n_F} (N_P, N_F) < \int_0^1 s w(s) f'_{n_F} (N_P, s N_F) \mathrm{d}s \\
       \text{or} \\
       f''_{n_F} (N_P, N_F) < \int_0^1 s^2 w(s) f''_{n_F} (N_P, s N_F) \mathrm{d}s.
   \end{gather*}
   $$
     - Holds for certain interesting cases, such as the example application
:::

## Equilibrium comparative statics

::::: {.columns}

:::: {.column width="40%"}

::: {.fragment fragment-index=1}
 - Hybrid mode reduces fringe entry
$$
\small
\frac{\partial N_F^*}{\partial N_P} < 0
$$
:::
::: {.fragment fragment-index=2}
 - And, under certain conditions, total profits
$$
\small\text{Let } f(N_P, N_F) = g(\alpha N_P + \beta N_F)
$$
$$
\small\text{Then } \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}
$$
:::

::::

:::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}

:::
::: {.fragment fragment-index=2}

:::

:::

::::

:::::

# More concrete example

# Summary

## References

::: {#refs}
:::