### 2023-04-20 src/presentation/topics_talk.qmd
Fix minor things

---
title: "Hybrid platforms and bargaining"
author: "Martin Stancsics"
format:
    revealjs:
        embed-resources: true
        html-math-method: mathjax
        transition: fade
        touch: true
        toc: true
        toc-depth: 1
        toc-title: Agenda
        slide-number: c
        progress: true
        controls: false
        theme:
         - default
         - css/custom.scss
        include-in-header:
         - js/mathjax-settings.html
bibliography: references.bib
---

# Introduction

<style>
    ul li li {
        font-size: 0.75em;
        line-height: 1.2em;
    }
</style>

## Motivation

::: {.highlight-block}
Examine the bargaining power implications of hybrid platforms and its welfare consequences
:::

::: {.fragment}
 - Hybrid platform
    - Getting more and more common
    - Seemingly obvious concerns and high-profile competition policy cases
    - Comparatively little reserach
:::
::: {.fragment}
 - Bargaining between participants is understudied in the platform setting
   - Authors generally assume take-it-or-leave-it offers
:::

::: {.notes}

:::

## Main ideas

::: {.fragment}
 - Shapley-value based bargaining rule
    - Analytically more tractable
    - Has bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
    - Has precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
:::
::: {.fragment}
 - Abstract away all frictions
    - Allows me to focus on the bargaining channel
    - Offers a good benchmark model
:::
::: {.fragment}
 - Focus on a continuum of small players case for tractability
    - Good approximation of the finite player case even for not too many players
    - Results have a nice geometric interpretation
:::

## Preview of results

::: {.fragment}
 - The impacts of hybrid platforms depend on the substitutability of the products
   - When they are substitutes, it increases the platform's profit share and decreases the fringe's profits
   - In the presence of investment costs, it can lead to lower fringe entry
   - With Logit demand, hybrid operation leads to welfare losses
:::
::: {.fragment}
 - This modeling strategy seems promising
   - The model is surprisingly tractable
   - Quite general results can be obtained
:::

## Related literature

::: {.incremental}
 - **Hybrid platform markets**<br>
 @anderson2021hybrid, @gutierrez2021welfare, @hagiu2022should
 - Similarly structured markets
   - Private labels in retail [@steiner2004nature]
   - Vertical integration [@hart1990vertical; @de2005vertical; @montez2007downstream]
 - Cooperative games in IO<br>
 @hart1990property, @levy1997individual, @inderst2003bargaining, @montez2007downstream

:::

# Model

## Players

::: {.fragment}
 - One **platform**: $P$
   - Without the platform, no value can be created
   - Might have its own products (hybrid mdoe)
:::
::: {.fragment}
 - A continuum of potential **fringe entrants**: $F_i, i \in \mathbb{R}_0^+$
   - Infinitesimally small
   - Have one product each
   - Can only sell through the platform
:::

## Production

Reduced-form approach: assume some function for **total profit**:

::: {.fragment}
$$
f(N_P, N_F)
$$
:::

::: {.fragment}
 - $N_P$ is the number of platform products
 - $N_F$ is the number of fringe products/firms
 - Assume that $f$ is increasing in both arguments
:::

## Profit sharing

::: {.fragment}
 - Platform gets $$ \pi_P(N_P, N_F) = \int_0^1 w(s) f(N_P, sN_F) \mathrm{d}s $$
 - Fringe gets the rest
:::
::: {.fragment}
 - Based on the Shapley-value 
   - Bargaining interpretation [e.g. @hart1996bargaining; @gul1989bargaining]
   - Precedents in the IO literature [e.g. @hart1990property; @montez2007downstream]
 - $w(s)$ is a weighting function with $\int_0^1 w(s) \mathrm{d}s = 1$
   - Related to the weighted value for a certain class of functions
:::

## Some simple comparative statics

::::: {.columns}

:::: {.column width="40%"}

::: {.fragment fragment-index=2}
 - The platform likes being hybrid
$$
\small
\frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0
$$
:::
::: {.fragment fragment-index=4}
 - It depends™ for the fringe
$$
\begin{align*}
\small\mathrm{sign} \left( \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} \right) = \\
\small\mathrm{sign} \left( \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} \right)
\end{align*}
$$
:::

::::

:::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_n_p_on_shares/800p60/sections/draw_graph.mp4" />
:::
::: {.fragment fragment-index=2}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_n_p_on_shares/800p60/sections/cross_derivative_positive.mp4" />
:::
::: {.fragment fragment-index=3}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_n_p_on_shares/800p60/sections/back_to_original.mp4" />
:::
::: {.fragment fragment-index=4}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_n_p_on_shares/800p60/sections/cross_derivative_negative.mp4" />
:::

:::

::::

:::::

## Entry and equilibrium

Now let us endogeneize the number of fringe entrants!

::: {.fragment}
 - Unlimited number of potential entrants
:::
::: {.incremental}
 - Timing
    - Hybrid mode is exogeneously determined $\to N_P$
    - Each fringe firm decides whether to invest at cost $I_F$ $\to N_F$
    - Platform and firms produce and sell products $\to f(N_P, N_F)$
    - Profits are shared according to previous rule $\to \pi_P(N_P, N_F), \pi_F(N_P, N_F)$
:::

## Equilibrium

::: {.incremental}
 - Free entry equilibrium: fringe firms make zero net profits
   $$ \pi_F(N_P, N_F^*) = I_F N_F^* $$
 - In order to guarantee uniqueness, let us assume
   $$
   \scriptsize
   \begin{gather*}
       f'_{n_F} (N_P, N_F) < \int_0^1 s w(s) f'_{n_F} (N_P, s N_F) \mathrm{d}s \\
       \text{or} \\
       f''_{n_F} (N_P, N_F) < \int_0^1 s^2 w(s) f''_{n_F} (N_P, s N_F) \mathrm{d}s.
   \end{gather*}
   $$
     - Holds for certain interesting cases, such as the example application
:::

## Equilibrium comparative statics

::::: {.columns}

:::: {.column width="40%"}

::: {.fragment fragment-index=3}
 - If $\frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} < 0$, hybrid mode reduces fringe entry

:::
::: {.fragment fragment-index=4}
 - And sometimes total profits, too
 $$
 \begin{gather*}
   g(\alpha N_P + \beta N_F), \\
   g'' < 0, w(s) \equiv 1
 \end{gather*}
 $$
 $$
 \implies \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}
 $$
:::

::::

:::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_equilibrium_entry/800p60/sections/draw_graph.mp4" />
:::
::: {.fragment fragment-index=2}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_equilibrium_entry/800p60/sections/mark_equilibrium.mp4" />
:::
::: {.fragment fragment-index=3}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_equilibrium_entry/800p60/sections/alternate_equilibrium.mp4" />
:::
::: {.fragment fragment-index=4}
  <video data-autoplay src="../../out/manim_figures/videos/comparative_equilibrium_entry/800p60/sections/show_loss.mp4" />
:::

:::

::::

:::::

# Microfounding the profit function

## Demand

::: {.incremental}
 - Follow @anderson2021hybrid
 - Unit mass of customers, each choosing one product maximizing $$u_{ij}^T = v^T - p_i^T + \mu \varepsilon_{ij}^T$$
   - $T \in \{P, F, 0\}$
   - Unit mass of outside options at price $p_i^0 = 0$
   - $v_T$: value of the product
   - $\mu$: degree of horizontal differentiation
   - $\varepsilon_{ij}^T \sim \mathrm{Gumbel}(0, 1)$: taste shocks
:::

## Demand

::: {.fragment}
Logit-like demand:
$$
x_{Ti} = \frac{\exp\left( \frac{v_T - p_{Ti}}{\mu} \right)}{A}
$$
:::
::: {.fragment}
where
$$
A = \int_0^{N_F} \exp\left( \frac{v_F - p_{Fi}}{\mu} \right) \mathrm{d}i + \int_0^{N_P} \exp\left( \frac{v_P - p_{Pi}}{\mu} \right) \mathrm{d}i + 1
$$
:::

## Production
::: {.fragment}
 - Assume that the platform prices its products as if they were made by seperate, competitive sellers
   - Possible interpretation: profit-maximizing subsidiaries
   - More importantly: "best case scenario"
:::
::: {.fragment}
 - The optimal price is an additive markup rule:
 $$
 \begin{align*}
 p^*_{Ti} &= c_T + \mu \\
 \pi^{v*}_{Ti} &= \mu \frac{\exp \left( \frac{v_T - c_T - \mu}{\mu} \right)}{A}
 \end{align*}
 $$
:::

## Entry and equilibrium

I consider two cases:

::: {.incremental}
 - **Benchmark**<br>platform can commit to any lump-sum entry fee $F_F$
 - **Bargaining**<br>total profits are shared according to the Shapley-value-based rule
:::

## Benchmark equilibrium

::: {.incremental}
 - Optimal entry fee is independent of $N_P$:
   $$ F_F^* = \sqrt{\mu I_F V_F} - I_F $$
 - Size of the aggregate is independent* of $N_P$:
   $$ A^* = \sqrt{\frac{\mu V_F}{I_F}} $$
 - **Consumer surplus** is also **independent*** of $N_P$:
   $$ CS^* \propto \log(A^*) $$
:::

## Bargaining equilibrium

::: {.incremental}
 - Total profits as a function of entrants have a simple form:
 $$
 f(N_P, N_F) = \mu \frac{N_F V_F + N_P V_P}{N_F V_F + N_P V_P + 1}
 $$
   - Satisfies the assumption for unique equilibrium
   - Has the form $f(N_P, N_F) = g(\alpha N_P + \beta N_F)$
   - $g$ is concave
 - Therefore,
 $$
 \frac{\partial N_F^*}{\partial N_P} < -\frac{V_P}{V_F}
 $$
:::

## Bargaining equilibrium

::: {.incremental}
 - Equilibrium aggregate and total profits are decreasing in $N_P$
 $$
 \frac{\partial A^*}{\partial N_P} < 0
 $$
   - Platform profits are increasing in $N_P$, though
 - **Consumer surplus** is also **decreasing** in $N_P$
 $$
 \frac{\partial CS^*}{\partial N_P} < 0
 $$
:::

## Results graphically

::: {.r-stack}

::: {.fragment fragment-index=1}
  <video height="500" data-autoplay src="../../out/manim_figures/videos/equilibrium_outcomes/800p60/sections/draw_graph.mp4" />
:::
::: {.fragment fragment-index=2}
  <video height="500" data-autoplay src="../../out/manim_figures/videos/equilibrium_outcomes/800p60/sections/a_nof.mp4" />
:::
::: {.fragment fragment-index=3}
  <video height="500" data-autoplay src="../../out/manim_figures/videos/equilibrium_outcomes/800p60/sections/a_bench.mp4" />
:::
::: {.fragment fragment-index=4}
  <video height="500" data-autoplay src="../../out/manim_figures/videos/equilibrium_outcomes/800p60/sections/a_barg.mp4" />
:::
::: {.fragment fragment-index=5}
  <video height="500" data-autoplay src="../../out/manim_figures/videos/equilibrium_outcomes/800p60/sections/pi_p_all.mp4" />
:::

:::

# Conclusion

## Summary

::: {.fragment}
 - **Tractable model** of hybrid platforms in which **bargaining** plays a key role
   - Applicable to other settings, too (e.g. vertical markets, franchises)
:::
::: {.fragment}
 - Highlight an important aspect if hybrid platforms
   - **Hybridness increases bargaining power** against entrants
   - This can have **negative welfare consequences**
:::
::: {.fragment}
 - Policy implications for such markets
   - Amazon operating in hybrid mode
   - Microsoft acquiring Activision/Blizzard for $68bn
:::

## Future directions

::: {.fragment}
 - In this paper
   - Endogenenize hybridness ($N_P$)
   - Relax separate pricing assumption
   - Alternative benchmarks (e.g. exogeneous split)
:::
::: {.fragment}
 - Longer term
   - Endogeneize consumer entry (possibly with them engaging in the negotiation)
   - Apply this bargaining rule to other, similar settings
:::

# Thank you {.unnumbered .unlisted}

## References

::: {#refs}
:::

# Appendix

## The Shapley-value

Each player gets their expected marginal contribution

:::: {.columns}

::: {.column width="50%"}
::: {.r-stack}

::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/draw_objects.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P12.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_P21.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_2P1.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_1P2.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_12P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/value_21P.mp4" />
:::
::: {.fragment}
   <video data-autoplay src="../../out/manim_figures/videos/shapley_value_demo/800p60/sections/final.mp4" />
:::

:::
:::

::: {.column width="10%"}
:::

::: {.column width="40%"}
Coalition | Profits
----------|---------
$\{P\}$     |  0
$\{A_i\}$   | 1
$\{A_1, A_2\}$  | 2
$\{P, A_i\}$  | 3
$\{P, A_1, A_2\}$  | 6
:::

::::

::: {.notes}
 - Most direct definition of Shapley value
 - Closed-form expression
 - Generally computationally expensive, but not in this case
:::

## Baseline model

:::: {.columns}

::: {.column width="100%"}

Set of all players: $N = \{P, A_1, \dots, A_n\}$

**Assumption**: the platform is the only way for firms to
reach users and that all $n$ firms are identical
$$
v(C) = \begin{cases}
   0                                           & \text{if } P \notin C \\
   f\left(\frac{|C \setminus \{P\}|}{n}\right) & \text{otherwise}.
\end{cases}
$$
$\forall\, C \subset N$.
:::

::::

::: {.notes}
 - Not only platforms. Also describes
   - upstream-downstream setups
   - generally any setting with an important large players and many identical small ones
:::

## Value of the platform

:::: {.columns}

::: {.column width="40%"}

Characteristic function:
$$
\scriptsize
v(C) = \begin{cases}
   0 \;\text{ if } P \notin C \\
   f\left(\frac{|C \setminus \{P\}|}{n}\right)
\end{cases}
$$

::: {.fragment .fade-in fragment-index=2}
Shapley value:
$$l 
\scriptsize
\varphi_P^n(f) = \frac{1}{n+1} \sum_{k=0}^{n} f(k)
$$
:::
::: {.fragment .fade-in fragment-index=5}
$$
\scriptsize
\to \int_0^1 f(t) \, \mathrm{d}t
$$
:::

:::

::: {.column width="60%"}

::: {.r-stack}

::: {.fragment fragment-index=1}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/draw_bars.mp4" />
:::
::: {.fragment fragment-index=2}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/add_brace.mp4" />
:::
::: {.fragment fragment-index=3}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/more_firms.mp4" />
:::
::: {.fragment fragment-index=4}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/continuous_approximation.mp4" />
:::
::: {.fragment fragment-index=6}
   <video data-autoplay src="../../out/manim_figures/videos/proposition_main/800p60/sections/surplus_division.mp4" />
:::

:::

:::

::::

::: {.notes}
 - Think of the previous figure with the orderings
 - Main insight: each ordering with the platform at position $k$ has the same probability
:::